---
title: "Oregon CEP Participation"
author: "Marisa Carter"
date: "2026-02-15"
output:
  pdf_document:
    latex_engine: xelatex
---

```{r}
library(tidyverse)
library(fixest)
library(did)
library(bacondecomp)
library(haven)
library(dplyr)
library(tidyverse)
library(janitor)
library(stringr)
library(tidyr)
library(readr)
library(pdftools)
library(knitr)
library(here)

cep <- read_csv(
  here("data", "processed", "OR_CEP_clean.csv"),
  show_col_types = FALSE
) %>%
  group_by(agency_name) %>%
  filter(
    all(total_students >= 50, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    treated = if_else(
      !is.na(first_treat_year) & year >= first_treat_year,
      1,
      0
    ),
    event_time = year - first_treat_year,
    first_treat_year_cs = if_else(
      is.na(first_treat_year),
      0,
      first_treat_year
    ),
    district_id = as.numeric(factor(agency_name))
  )

```

# Check that panel is set up correctly
```{r}
cep_check <- cep %>%
  select(
    agency_name,
    district_id,
    school_year,
    year,
    first_treat_year,
    first_treat_year_cs,
    CEP_treat,
    treated,
    event_time
  ) %>%
  arrange(agency_name, year)

View(cep_check)

# Number of districts
cep %>%
  summarise(
    n_districts = n_distinct(agency_name)
  )

# Cohort sizes
cep %>%
  distinct(agency_name, first_treat_year_cs) %>%
  count(first_treat_year_cs)

# Event-time support check
cep %>%
  filter(treated == 1) %>%
  distinct(agency_name, event_time) %>%
  count(event_time) %>%
  arrange(event_time)
```

# Summary statistics
```{r}
## Number of districts, years, observations
cep %>%
  summarise(
    districts = n_distinct(agency_name),
    years = n_distinct(year),
    observations = n()
  )

### CEP adoption over time
cep %>%
  distinct(agency_name, first_treat_year) %>%
  mutate(adopt_year = ifelse(is.na(first_treat_year), "Never", first_treat_year)) %>%
  count(adopt_year)

### treatment rate by year

cep %>%
  group_by(year) %>%
  summarise(
    districts = n_distinct(agency_name),
    treated = sum(treated),
    share_treated = mean(treated)
  )

```

## Revenue, enrollment, poverty (MEPS)
```{r}
### unweighted by enrollment
cep %>%
  summarise(
    mean_total_rev = mean(total_rev, na.rm = TRUE),
    sd_total_rev = sd(total_rev_per_pupil, na.rm = TRUE)
  )
cep %>%
  summarise(
    mean_federal = mean(total_rev_fed, na.rm = TRUE),
    mean_state = mean(total_rev_state, na.rm = TRUE),
    mean_local = mean(total_rev_local, na.rm = TRUE)
  )

cep %>%
  summarise(
    mean_enrollment = mean(total_students, na.rm = TRUE),
    median_enrollment = median(total_students, na.rm = TRUE)
  )
```

## Revenue composition
```{r}
cep <- cep %>%
  mutate(
    federal_share = total_rev_fed /
      total_rev_per_pupil,

    state_share = total_rev_state /
      total_rev_per_pupil,

    local_share = total_rev_local /
      total_rev_per_pupil
  )

cep %>%
  summarise(
    mean_federal_share = mean(federal_share, na.rm = TRUE),
    mean_state_share = mean(state_share, na.rm = TRUE),
    mean_local_share = mean(local_share, na.rm = TRUE)
  )

### revenue over time
cep %>%
  group_by(year) %>%
  summarise(
    federal = mean(federal_share, na.rm = TRUE),
    state = mean(state_share, na.rm = TRUE),
    local = mean(local_share, na.rm = TRUE)
  )

cep %>%
  group_by(year) %>%
  summarise(
    federal = mean(federal_share, na.rm = TRUE),
    state = mean(state_share, na.rm = TRUE),
    local = mean(local_share, na.rm = TRUE)
  ) %>%
  pivot_longer(-year, names_to = "source", values_to = "share") %>%
  ggplot(aes(year, share, color = source)) +
  geom_line(size = 1.2) +
  geom_point() +
  labs(
    y = "Share of Revenue",
    x = "Year",
    title = "Revenue Composition of Oregon School District Funding"
  ) +
  theme_minimal()


### revenue composition by CEP status
cep %>%
  group_by(treated) %>%
  summarise(
    federal_share = mean(federal_share, na.rm = TRUE),
    state_share = mean(state_share, na.rm = TRUE),
    local_share = mean(local_share, na.rm = TRUE)
  )

### pre-treatment revenue composition
cep %>%
  filter(year < first_treat_year | is.na(first_treat_year)) %>%
  mutate(eventually_treated = if_else(is.na(first_treat_year), 0, 1)) %>%
  group_by(eventually_treated) %>%
  summarise(
    federal_share = mean(federal_share, na.rm = TRUE),
    state_share = mean(state_share, na.rm = TRUE),
    local_share = mean(local_share, na.rm = TRUE)
  )

```

# Baseline TWFE
## total rev
```{r}
model_total_twfe <- feols(
  total_rev_per_pupil ~ treated |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

summary(model_total_twfe)
```

## Federal, state, local revenue
```{r}

model_federal_twfe <- feols(
  total_rev_fed ~ treated |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

model_state_twfe <- feols(
  total_rev_state ~ treated |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

model_local_twfe <- feols(
  total_rev_local ~ treated |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

summary(model_federal_twfe)
summary(model_state_twfe)
summary(model_local_twfe)
```

# TWFE event study
```{r}
model_event_twfe <- feols(
  total_rev_per_pupil ~
    i(event_time, ref = -1) |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

summary(model_event_twfe)
iplot(model_event_twfe)

model_total_twfe_controls <- feols(
  total_rev_per_pupil ~
    treated + share_child_poverty |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

summary(model_total_twfe_controls)
```

# federal, state, local rev
```{r}
model_event_federal <- feols(
  total_rev_fed ~
    i(event_time, ref = -1) |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

summary(model_event_federal)
iplot(model_event_federal)


model_event_state <- feols(
  total_rev_state ~
    i(event_time, ref = -1) |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

summary(model_event_state)
iplot(model_event_state)


model_event_local <- feols(
  total_rev_local ~
    i(event_time, ref = -1) |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

summary(model_event_local)
iplot(model_event_local)





etable(model_event_federal, model_event_state, model_event_local)

cep <- cep %>%
  mutate(event_time_bin = case_when(
    event_time <= -4 ~ -4,
    event_time >= 4 ~ 4,
    TRUE ~ event_time
  ))

```


# Callaway-Sant'anna DiD
## Total revenue
```{r}
table(cep$first_treat_year_cs)
### not yet treated districts.

cs_model <- att_gt(
  yname = "total_rev_per_pupil",
  tname = "year",
  idname = "district_id",
  gname = "first_treat_year_cs",
  data = cep,
  panel = TRUE,
  clustervars = "district_id"
)

summary(cs_model)

# Overall average ATT
cs_overall <- aggte(
  cs_model,
  type = "simple"
)

summary(cs_overall)

# Dynamic / event-study ATT
cs_dynamic <- aggte(
  cs_model,
  type = "dynamic"
)

summary(cs_dynamic)
ggdid(cs_dynamic)


# export

cs_dynamic_df <- tibble(
  event_time = cs_dynamic$egt,
  att = cs_dynamic$att.egt,
  se = cs_dynamic$se.egt
) %>%
  mutate(
    ci_low = att - 1.96 * se,
    ci_high = att + 1.96 * se
  )

write_csv(
  cs_dynamic_df,
 "/Users/marisalenaycarter/Desktop/CEP-Fiscal-Effects/output/cs_dynamic_effects_totalrev_unweighted.csv"
)


```

## Total revenue, weighted by enrollment
```{r}
cep <- cep %>%
  filter(year <= 2020)

names(cep)

t_model <- att_gt(
  yname = "total_rev_per_pupil",
  tname = "year",
  idname = "district_id",
  gname = "first_treat_year_cs",
  data = cep,
  panel = TRUE,
  clustervars = "district_id",
  weightsname = "total_students"
)

summary(t_model)

# Overall average ATT
t_overall <- aggte(
  t_model,
  type = "simple"
)

summary(t_overall)

# Dynamic / event-study ATT (truncated window)
t_dynamic <- aggte(
  t_model,
  type = "dynamic",
  min_e = -4,
  max_e = 5
)

summary(t_dynamic)
ggdid(t_dynamic)

# export
t_dynamic_df <- tibble(
  event_time = t_dynamic$egt,
  att = t_dynamic$att.egt,
  se = t_dynamic$se.egt
) %>%
  mutate(
    ci_low = att - 1.96 * se,
    ci_high = att + 1.96 * se
  )

write_csv(
  t_dynamic_df,
 "/Users/marisalenaycarter/Desktop/CEP-Fiscal-Effects/output/cs_dynamic_effects_totalrev.csv"
)

summary(t_overall)
```

## Federal revenue weighted by enrollment
```{r}
cep <- cep %>%
  filter(year <= 2020)

names(cep)

f_model <- att_gt(
  yname = "total_rev_fed",
  tname = "year",
  idname = "district_id",
  gname = "first_treat_year_cs",
  data = cep,
  panel = TRUE,
  clustervars = "district_id",
  weightsname = "total_students"
)

summary(f_model)

# Overall average ATT
f_overall <- aggte(
  f_model,
  type = "simple"
)

summary(f_overall)

# Dynamic / event-study ATT (truncated window)
f_dynamic <- aggte(
  f_model,
  type = "dynamic",
  min_e = -4,
  max_e = 4
)

summary(f_dynamic)
ggdid(f_dynamic)

# export
f_dynamic_df <- tibble(
  event_time = f_dynamic$egt,
  att = f_dynamic$att.egt,
  se = f_dynamic$se.egt
) %>%
  mutate(
    ci_low = att - 1.96 * se,
    ci_high = att + 1.96 * se
  )

write_csv(
  f_dynamic_df,
  "/Users/marisalenaycarter/Desktop/CEP-Fiscal-Effects/output/cs_dynamic_effects_federalrev.csv"
)

summary(f_overall)
```

### State revenue weighted by enrollment
```{r}
cep <- cep %>%
  filter(year <= 2020)

names(cep)
s_model <- att_gt(
  yname = "total_rev_state",
  tname = "year",
  idname = "district_id",
  gname = "first_treat_year_cs",
  data = cep,
  panel = TRUE,
  clustervars = "district_id",
  weightsname = "total_students"
)

summary(s_model)

# Overall average ATT
s_overall <- aggte(
  s_model,
  type = "simple"
)

summary(s_overall)

# Dynamic / event-study ATT (truncated window)
s_dynamic <- aggte(
  s_model,
  type = "dynamic",
  min_e = -4,
  max_e = 5
)

summary(s_dynamic)
ggdid(s_dynamic)

# export
s_dynamic_df <- tibble(
  event_time = s_dynamic$egt,
  att = s_dynamic$att.egt,
  se = s_dynamic$se.egt
) %>%
  mutate(
    ci_low = att - 1.96 * se,
    ci_high = att + 1.96 * se
  )

write_csv(
  s_dynamic_df,
 "/Users/marisalenaycarter/Desktop/CEP-Fiscal-Effects/output/cs_dynamic_effects_staterev.csv"
) 

summary(s_overall)
```


# Sun & Abraham
## Total revenue
```{r}
model_sunab <- feols(
  total_rev_per_pupil ~
    sunab(first_treat_year, year) |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

summary(model_sunab)

iplot(model_sunab, xlim = c(-5, 5))


range(cep$year, na.rm = TRUE)
sort(unique(cep$year))
sort(unique(cep$first_treat_year))
```

## Federal revenue
```{r}
model_federal_sunab <- feols(
  total_rev_fed ~
    sunab(first_treat_year, year) |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

iplot(model_federal_sunab, xlim = c(-5,5))
```

## State revenue
```{r}
model_state_sunab <- feols(
  total_rev_state ~
    sunab(first_treat_year, year) |
    agency_name + year,
  cluster = ~agency_name,
  data = cep
)

iplot(model_state_sunab, xlim = c(-5,5))

```


# Goodman-Bacon
## Total Revenue
```{r}
cep_bacon <- cep %>%
  select(
    agency_name,
    year,
    treated,
    total_rev_per_pupil
  ) %>%
  filter(
    !is.na(total_rev_per_pupil),
    !is.na(treated),
    !is.na(year),
    !is.na(agency_name)
  )

balanced_ids <- cep_bacon %>%
  count(agency_name) %>%
  filter(n == length(unique(cep_bacon$year))) %>%
  pull(agency_name)

cep_bacon <- cep_bacon %>%
  filter(agency_name %in% balanced_ids)

bacon_results <- bacon(
  total_rev_per_pupil ~ treated,
  data = cep_bacon,
  id_var = "agency_name",
  time_var = "year"
)

bacon_results
```


